---
title: "R Assignment 2 Notebook"
output: html_notebook
# Analysis of LendingClub.com Data
# Dr. Clif Baldwin
# I downloaded data from LendingClub.com  July 25, 2019  
# Modified by Daniel W. Anner
# Modified on 10/26/2020
---
# Load any libraries you need
```{r}
library(tidyverse)
library(tibble)
library(readr)
library(dplyr)
library(lubridate)
setwd('/home/rstudio/dssa_workspace/R/Assignment2')
```

# Read in the data
```{r}
loan <- read_csv("loans.csv")
# or 
#loan <- read.csv("loans.csv")
```

# Look at the data
```{r}
glimpse(loan)
unique(loan$loan_status)
```

# Detect who is late?
# In other words, what observations have the word "Late" in the loan_status field?
```{r}
#loan %>% filter_at(vars(loan_status), any_vars(. %in% 'Late'))
loan %>% filter(loan_status == "Late (31-120 days)" | loan_status == "Late (16-30 days)")
```
#We can see there are 218 people with late loans (including all, 16-120 days)

# Define a new variable (or column) to indicate whether the loan was late (0) or not late (1)
```{r}
loan_latestatus <- loan
loan_latestatus$late_state = 0
loan_latestatus <- mutate(loan_latestatus, late_state = replace(late_state, loan_status == "Late (31-120 days)", 1))
loan_latestatus <- mutate(loan_latestatus, late_state = replace(late_state, loan_status == "Late (16-30 days)", 1))
```

# Calculate the correlation between late and annual income 
#   Hint: use cor()
#   Note: the parameter use="complete.obs" ignores "NA"
```{r}
loan_latestatus <- transform(loan_latestatus, loan_state = as.numeric(late_state))
cor(loan_latestatus$loan_state, loan_latestatus$annual_inc, use="complete.obs")
```
#Very small almost non-existant coorelation


# The emp_length field is character with some "n/a"
# Extract just the numbers and change "n/a" to NA (hint: use parse_number())
```{r}
loan_latestatus$emp_length <- parse_number(loan_latestatus$emp_length, na= c("n/a", "NA"))
```

# Compute the correlation of late and years employed
```{r}
cor(loan_latestatus$late_state, loan_latestatus$emp_length, use="complete.obs")
```
#Interesting, a slightly negative correlation

# Plot a histogram of annual income 
```{r}
ggplot(data=loan_latestatus, aes(x=annual_inc)) +
  geom_histogram(binwidth=10000) +
  xlim(0, max(loan_latestatus$annual_inc))
```

#A much better graph looking histogram removing the top most earners
```{r}
ggplot(data=loan_latestatus, aes(x=annual_inc)) +
  geom_histogram(binwidth=10000) +
  xlim(0,500000)
```

# Compute the summary statistics of annual income
```{r}
summary(loan_latestatus$annual_inc)
```

# Explore the date columns
# You will notice they are in month-year format.
# You can convert the string month-year to a date with 
```{r}
loan_latestatus$issue_d <-  parse_date_time(loan_latestatus$issue_d, "my")
loan_latestatus$last_pymnt_d <-  parse_date_time(loan_latestatus$last_pymnt_d, "my")
loan_latestatus$next_pymnt_d <-  parse_date_time(loan_latestatus$next_pymnt_d, "my")
loan_latestatus$last_credit_pull_d <-  parse_date_time(loan_latestatus$last_credit_pull_d, "my")
```

# Do something with this datetime and just explore the data
```{r}
loan_latestatus$overdue_time <- difftime(loan_latestatus$next_pymnt_d, loan_latestatus$last_pymnt_d, units="days")
loan_latestatus[ , c("loan_amnt", "next_pymnt_d", "last_pymnt_d", "overdue_time")]
```
#Made a simple table and added an overdue time column
```{r}
ggplot(data=loan_latestatus, aes(x=overdue_time)) +
  geom_histogram(breaks=seq(0,100,5))
```
#Also made a histogram of the overdue times

```{r}
summary(loan_latestatus$issue_d)
```






























