---
title: "R Notebook"
output: html_notebook
---

First we setup the libraries and the data variable, where we use read_csv to import our CSV to a machine readable format
```{r}
library(readr)
library(tibble)

setwd('/home/rstudio/dssa_workspace/R/Assignment1')
data = read_csv('fish.csv')
```
<hr />
<h2>Lets head() the data</h2>

```{r}
head(data)
```

<h4>What did we learn?</h4>
Well we can see our rows and columns which show us yearly data for counties, with brown trout sizes.

<hr />
<h2>Lets dim() the data</h2>

```{r}
dim(data)
```

<h4>What did we learn?</h4>
Well we can see that there are 5,580 rows and 8 columns.

<hr />
<h2>Lets nrow() the data</h2>

```{r}
nrow(data)
```

<h4>What did we learn?</h4>
Well we can see that there are 5,580 rows just as dim shows us.

<hr />
<h2>Lets glimpse() the data</h2>

```{r}
glimpse(data)
```

<h4>What did we learn?</h4>
Well we can see some more information in a different view which shows us there are 5,580 rows (just as nrow and dim showed us), as well as a little bit of data and row data types.

<hr />
<h2>Lets summary() the data</h2>

```{r}
summary(data)
```

<h4>What did we learn?</h4>
We can see some more information about each of the fields including information about the min and max data for the years, total number and sizes.

<hr />
<h2>Lets str() the data</h2>

```{r}
str(data)
```

<h4>What did we learn?</h4>
Str shows us a bit more information about the column types as well as some of the data. 
