---
title: "Linear Regression - DSSA 5201"
owner: "Daniel W. Anner"
output: html_notebook
---

# PseudoCode
# Read in the data
# Divide data into training and test data
# Train the model using the dataset:
#  Create X Matrix [1, xi1, xi2,...,xik] from all rows i of training data and for k variables
#  Create Y Matrix (vector)
# Solve for the β vector, 
# Make predictions using the test dataset:
#  Create X Matrix [1, xi1, xi2,...,xik] from all rows i of test data and for k variables
#  Compute Ŷ = βX
# Test the Model:
#  Compute the Sum of Squared Errors RSS 
#  Compute the Total Sum of Squares TSS 
#  Compute R2 = 1 – RSS / TSS 

```{r}
setwd("C:/Users/danie/_Source/Stockton-DSSA/DSSA-5201_Machine_Learning_Fundamentals/LinearRegression")
# Get the data
df <- read.csv("TrainData_Group1.csv", header = TRUE)
```

```{r}
# Separate data into test and train. Train will be most of data set and test will be the remaining
indexes <- sample(1:nrow(df), size=0.80*nrow(df)) 
# Set data aside for training
train <- df[indexes,]
# Set data aside for testing
test <- df[-indexes,]
# Remove the index variable from memory - helps with larger data sets - no need to retain in memory data we do not need
rm(indexes)
```

```{r}
# Linear regression function
Linear_Regression <- function(df) {

}
```

```{r}
# Y = βX
# Make predictions using test data
y_hat <- function(df, B_hat) {
   
}
```

```{r}
#SSE TSS and R^2
ER <- function(y, y_hat, df) {

}
```

```{r}
# Linear regression test function
Linear_Regression_Testing <- function(train, test){

}
```

```{r}
# Run linear regression on training data 

```

